---
title: "kz2383_assignment2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## problem 2 


read and clean sheet mr_trash_wheel
```{r}
library(tidyverse)
library(readxl)
```
```{r}
 mr_trash_wheel = 
  read_excel(
    "~/Dropbox/Mac/Downloads/Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster)%>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls))%>%
  mutate(dumpster=as.character(dumpster),
         trash_wheel=c("mr"))
  

  
```


read and clean professor trash wheel data
```{r}
professor_trash_wheel = 
  read_excel(
    "~/Dropbox/Mac/Downloads/Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel", range = cell_cols("A:M")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster)%>%
  mutate(
         dumpster=as.character(dumpster),
         year=as.character(year),
          trash_wheel=c("prof"))




```

combine mr. trash wheel and professor trash wheel data

```{r}
data=mr_trash_wheel%>%full_join(professor_trash_wheel)


#total weight in dataset professor trash wheel

total_weight=sum(professor_trash_wheel$weight_tons)
total_weight

# total number of sports balls in 2020 in mr. trash wheel 
num=mr_trash_wheel%>%filter(year=="2020")%>%pull(sports_balls)
total_ball=sum(num)
total_ball
 

 
```
 For Mr. Trash Wheel there are  `r nrow(mr_trash_wheel)` observations and `r ncol(mr_trash_wheel)` variables. Except for  `month` and `date`, all variables are  `integer` or `numeric` variables. Number of sports balls is rounded to nearest integer so is the only integer variable in the dataset.
 
 As for professor trash wheel, there are `r nrow(professor_trash_wheel)` observations and `r ncol(professor_trash_wheel)` variables.Except for  `month` and `date`, all variables are  `integer` or `numeric` variables. Number of sports balls is rounded to nearest integer so is the only integer variable in the dataset.
 
 ### There are a total of  tons of trash collected by professor trash wheel and  sports balls collected by mr. trash wheel in 2020. 
 
 
 
 
## problem 3 


clean pols-month dataset, separate variable mon and create new variable president

```{r}
pols = 
  read_csv("/Users/zhankeming/Desktop/zhank/Desktop/data science/solution for data science/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>%
  mutate(
    month = month.name[month], 
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) %>%
  select(-day, -starts_with("prez"))

pols
```

 
clean dataset snp and arrange year and month as leading columns

```{r}
snp = 
  read_csv("/Users/zhankeming/Desktop/zhank/Desktop/data science/solution for data science/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("month", "day", "year"), convert = TRUE) %>%
  arrange(year, month) %>%
  mutate(month = month.name[month]) %>%
  select(year, month, close) 

snp
```
transform unemployment data from wide to long format

```{r}
unemployment = 
  read_csv("/Users/zhankeming/Desktop/zhank/Desktop/data science/solution for data science/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/unemployment.csv") %>%
  gather(key = month, value = unemployment, Jan:Dec) %>%
  mutate(
    month = 
      recode(month, 
             Jan = "January", Feb = "February",
             Mar = "March", Apr = "April",
             May = "May", Jun = "June",
             Jul = "July", Aug = "August",
             Sep = "September", Oct = "October",
             Nov = "November", Dec = "December"))

unemployment 
```

merge pols-month, snp and unemployment datasets

```{r}
data = 
  left_join(pols, snp) %>%
  left_join(., unemployment)

data
```


The first dataset `pols` includes the number of national politicians who are democratic or republican at any given time.

The second dataset `snp` indicates the stock market index (S&P), which usually represents the situation of stock market as a whole.
 
The third dataset `unemployment` showed the percentage of unemployment in each month of the associated year.

In the final merged dataset `data`,

* there are `r nrow(data)` rows and `r ncol(data)` columns;
* it range from year `r min(as.numeric(pull(data,year)))` to year `r max(as.numeric(pull(data,year)))`;
* it have key variables named "`r names(data)[-c(1:2)]`".




